Bird beak's elaboration and exploration

### 2021-06-02 Analysis/03-elaboration_exploration_analyses.Rmd
analyses up

```{r, eval = TRUE}
set.seed(42)
## Selecting 1000 random covariance matrices from the MCMCglmm
covar_matrices <- get.covar(combined_results, n = 100)
```

@@

> The detailed procedure was adapted from [李哲源's post on Stack Overflow](https://stackoverflow.com/questions/40300217/obtain-vertices-of-the-ellipse-on-an-ellipse-covariance-plot-created-by-care/40316331#40316331) and implemented in `beer::get.axis` ([specific procedure on lines 40-73](https://github.com/TGuillerme/elaboration_exploration_bird_beaks/blob/master/beer/R/get.axes.R)).

```{r}
## Calculating the major axes
major_axes <- get.axes(covar_matrices, centre = "intercept")
```

@@

```{r}
## Plotting the ellipses
plot.ellipses(covar_matrices, col = c("grey",colour_vector),
              add = FALSE)
plot.axes(major_axes, col = c("grey",colour_vector),
              add = TRUE)
legend("topleft", legend = c("all_clades", levels(demo_data$clade)), col = c("grey", colour_vector), pch = 19)
```

@@

```{r, echo = FALSE, eval = TRUE}
source("../Functions/test.rotation.R")
```
Figure: in a space with 5 elements: A, B, C, D, E (in grey); where D and E represent the major axes (e.g. the major phylogenetic effect axes from the MCMCglmm), we can rotate and rescale each element so that D and E become the unit vector of length 1 (the black letters D and E) and get the exploration elaboration scores for either the elements (e.g. element A's projection in blue and its rejection in orange) or from any other axes (e.g. the major phylogenetic effect axes from the MCMCglmm from a specific clade).

@@

```{r}
## Get all the angles and stuff
group_results <- analyses.group(major_axes[1:4], base = "animal")
names(group_results) <- levels(demo_data$clade)
```

The results are reported as boxplots of **log values** for each group indicating which group explores/elaborates more (and to which degree - angle) relatively to each other.

```{r, fig.height = 18, fig.heigh = 6}
## Plot the blob wise metrics
plot.analyses.group(lapply(group_results, log), col = colour_vector)
```

@@

And here are the **log values** plotted by group:

```{r}
## Plot results
plot.analyses.tip(lapply(tip_results_global, log), col = colour_vector, group = split(rownames(demo_data), f = demo_data$clade))
```

>>>>

```{r, eval = TRUE}
set.seed(42)
## Selecting 1000 random covariance matrices from the MCMCglmm
covar_matrices <- get.covar(combined_results, n = 1000)[1:4]
```

@@

Finally, the centred the matrix $M_{1,m}$ on the centroid of each clade.

> The detailed procedure was adapted from [李哲源's post on Stack Overflow](https://stackoverflow.com/questions/40300217/obtain-vertices-of-the-ellipse-on-an-ellipse-covariance-plot-created-by-care/40316331#40316331) and implemented in `beer::get.axes` ([specific procedure on lines 40-73](https://github.com/TGuillerme/elaboration_exploration_bird_beaks/blob/master/beer/R/get.axes.R)).

```{r}
## Get the centre of each clade
group_centres <- list(
    "gulls"      = colMeans(demo_data[demo_data$clade == "gulls", c(1:4)]),
    "plovers"    = colMeans(demo_data[demo_data$clade == "plovers", c(1:4)]),
    "sandpipers" = colMeans(demo_data[demo_data$clade == "sandpipers", c(1:4)]),
    "phylogeny"  = colMeans(demo_data[, c(1:4)]))

## Calculating the major axes
major_axes <- get.axes(covar_matrices, centre = group_centres)
```

@@

```{r}
set.seed(51)
## Sub-selecting 5 random VCV matrices for readibility
sub_sample_matrices <- get.covar(combined_results, 5)[1:4]
sub_sample_major_axes <- get.axes(sub_sample_matrices, centre = "intercept")

## Plotting the ellipses
plot.ellipses(sub_sample_matrices, col = c("grey",colour_vector),
              add = FALSE, transparent.scale = 0.7)
plot.axes(sub_sample_major_axes, col = c("grey",colour_vector),
              add = TRUE, transparent.scale = 0.7)
legend("topleft", legend = c("phylogeny", levels(demo_data$clade)), col = c("grey", colour_vector), lty = 1)
```

We can also visualise all the major axes in relation to the elements in the space.

```{r, fig.height = 18, fig.width = 18}
## Plotting it
plot.space(demo_data,
           col = colour_vector,
           levels = demo_data$clade,
           xlab = "PC1 (90.5%)",
           ylab = "PC1 (6.86%)",
           xlim = c(-2, 2),
           ylim = c(-1, 1))
points(do.call(rbind,group_centres)[,c(1,2)],
       pch = 13, col = c(colour_vector, "grey"))
plot.axes(major_axes["animal"], col = "grey",
          add = TRUE, transparent.scale = 0.01,
          centre = 0)
plot.axes(major_axes["animal:clade1"], col = colour_vector[1],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "gulls", c(1,2)]))
plot.axes(major_axes["animal:clade2"], col = colour_vector[2],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "plovers", c(1,2)]))
plot.axes(major_axes["animal:clade3"], col = colour_vector[3],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "sandpipers", c(1,2)]))
legend("topleft", legend = c("phylogeny", levels(demo_data$clade)), col = c("grey", colour_vector), lty = c(1,1,1,1), pch = c(NA, 19, 19, 19))
```

@@

```{r, echo = FALSE, eval = TRUE}
source("../Functions/test.rotation.R")
```

Figure: in a space with 5 elements: A, B, C, D, E (in grey); where D and E represent the major axes (e.g. the major phylogenetic effect axes from the MCMCglmm), we can rotate and rescale each element so that D and E become the unit vector of length 1 (the black letters D and E) and get the exploration elaboration scores for either the elements (e.g. element A's projection in blue and its rejection in orange) or from any other axes (e.g. the major phylogenetic effect axes from the MCMCglmm from a specific clade).

@@

```{r}
## Get all the angles and stuff
group_results <- analyses.group(major_axes, base = "animal")
names(group_results) <- levels(demo_data$clade)
```

The results are reported as boxplots for each group indicating which group explores/elaborates more (and to which degree - angle) relatively to each other.

```{r, fig.height = 10, fig.heigh = 10}
## Plot the blob wise metrics
plot.analyses.group(group_results, col = colour_vector)
```

@@

And here are the values plotted by group:

```{r}
## Plot results
plot.analyses.tip(tip_results_global, col = colour_vector, group = split(rownames(demo_data), f = demo_data$clade))
```
